---
title: "Course7Project-YR"
author: "Yonatan Rafael"
output:
  html_document: default
  pdf_document: default
---
==============================================================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Overview

Part 1 of this report illustrates a simulation of the exponential distribution and it's proximity to the theoretical distribution, along with proof that the distribution is approximately normal.

### Sample versus Theoretical Mean

Run 1000 simulations of an exponential distribution. Each simulation contains 40 observations and a lamba = .2. The output returns the sample mean of our simulation.
```{r}
n <- 40
lambda <- .2
nosim <- 1000
df <- data.frame(ncol = 2, nrow = nosim)
colnames(df) <- c("index", "simulation")
set.seed(1000)
for (i in 1:nosim) {
        df[i, 1] <- i
        df[i, 2] <- mean(rexp(n, lambda))
}
mean(df$simulation)
```

The theoretical mean is calculated below.
```{r}
1/ lambda
```

Below is a visual representation of the simulation distribution, with lines showing the proximity between the sample mean (4.99) and theoretical mean (5).
```{r}
hist(df$simulation, xlab = "Sample Mean", main = "Histogram of Sample Means")
abline(v = mean(df$simulation), col = 2, lwd = 3)
abline(v = 1 / lambda, col = 3, lwd = 3)
legend('topright', c("Sample Mean", "Theoretical Mean"),  col = c(col = 2, col = 3), lty = c(1,1))
```



### Sample versus Theoretical Variance

Using the simulation presented above, below is the sample variance.
```{r}
var(df$simulation)
```

The theoretical variance is calculated below.
```{r}
(1/ (lambda ** 2))/n
```

Below is another visual representation of the simulation distribution, with lines showing the proximity between the sampel variance (.654) and theoretical variance (.625).
```{r}
hist(df$simulation, xlab = "Sample Mean", main = "Histogram of Sample Means", prob = TRUE)
lines(density(df$simulation),col = 2,lwd = 3)
xfit <- seq(min(df$simulation), max(df$simulation), length = 100)
yfit <- dnorm(xfit, mean = 1/lambda, sd = (1/lambda/sqrt(n)))
lines(xfit, yfit, col = 3, lty = 2)
legend('topright', c("Sample Mean", "Theoretical Mean"),  col = c(col = 2, col = 3), lty = c(1,1))
```

### Proving Normal Distribution

The chart above starts to show that the distribution resembles a normal distribution. That said, let's use a QQ-plot to prove normality.

```{r}
qqnorm(df$simulation, col = "red")
qqline(df$simulation)
```


